Swarm coordination through perimeter detection

### 2017-03-22 paper.tex
Fixed Internal References

\begin{table}
\begin{center}
\begin{tabular}{| p{1.5cm} | c | p{4cm} |}
\hline
\bf Weight \bf Component & \bf Hexagonal & \bf Description \\ \hline
Sample Rate & 100 & ms - Unit sampling interval\\  \hline
$k_c$ & 5 & weight adjuster for cohesion vector\\  \hline
$k_r$ & 15 & weight adjuster for repulsion  vector\\  \hline
$k_d$ & 0/100 & weight adjuster for destination vector 0 for static baseline 100 for destination based\\  \hline
Neighbour Distance & 70 & units\\  \hline
Agent Boundary & 80 & units\\  \hline
Speed & 20 & units/s\\  \hline
\end{tabular}\caption{Swarm Weighted Model} \label{tab:DynamicPhysics1}
\end{center}
\end{table}

@@

\textit{Assumption 1:} The direction of the swarm is based upon having a fixed goal that is known to all the agents (Section \ref{sec:Direction1}). \\
\textit{Assumption 2:} The goal is migrated towards based on a simulated GPS signal and the bias $k_d$ is applied to the \textit{destination vector} to influence the movement of the agents.

@@

\section{Identifying the coordinator role}
For an agent to change its role from a standard agent to a coordinator there needs to an algorithm that identifies a set of conditions to trigger the change. This same algorithm also triggers an agent to revert to being a standard agent. The coordinator selection algorithm is separate from the agent coordination algorithms discussed in chapter~\ref{chapter:methods}.

@@

\section{Monolithic swarm - (all-agent)} 
In a monolithic swarm \cite{BS:13} the propagation towards the goal is achieved by all the agents using a GPS signal to give each agents' movement a \textit{direction vector} towards a required goal~(Equation~\ref{eq:BotDirection1}). There is no selection criteria and no computational overhead for the agents.

@@

\section{Simple multifaceted swarm (basic-count)} 
In a simple multifaceted swarm the propagation towards the goal is achieved by generating a \textit{destination vector} for an agent when the agent conforms to a simple counting rule. This \textit{destination vector} is obtained by using the agent's GPS to identify its current coordinate and generating a vector from that coordinate to a specific destination point. This rule creates a subset of agents that apply a local \textit{direction vector} to their movement. This has the effect of creating a directional effect on the whole swarm. The aggregate \textit{direction vector} magnitude of the swarm is less than that produced when all agents apply a \textit{direction vector}. This reduction is further diluted by agents without the \textit{direction vector} calculating a \textit{movement vector} to an equilibrium position~(Equation~\ref{eq:BotDirection1}). This change in the aggregate \textit{direction vector} will impact on a swarm's `jitter'. This is discussed in more detail in~Section~\ref{sec:AlternateBias1}. With there only being a limited number of peripheral agents invoking their GPS signal there will also be a net reduction in the energy usage of the swarm~(Figure~\ref{reduced:EnergyReductionGPS}).

A swarm with the appropriate field effects will propagate towards hexagons made from a central agent and six neighbours~(Figure~\ref{fig:StableForms}). A swarm can be influenced by its surroundings, which involves the interaction of other agents (as neighbours) and obstacles which can cause a compression of the agents. Compression can also occur during the stabilisation phase of a swarm. These situations can cause an agent to have less than 6 neighbours but still be surrounded. A general rule therefore to detect a rudimentary boundary of a hexagonal swarm is to calculate the number of neighbours an agent has. If the total is less than five ($nbr(b) < 5$) then there is a high probability that the agent is either in a void within the swarm or on an external perimeter. If an agent is on a perimeter or a void the it should enable its GPS and become a coordinator and provide a directional bias to the swarm via its \textit{direction vector}. This coordinator role affects the immediate neighbours \textit{movement vector} through cohesion/repulsion~(Figure~\ref{reduced:SimpleMulti12}). This simple detection algorithm allows a directional bias to be induced into a swarm with minimal computational impact on the coordinator agents.

@@

\subsection{Basic-count effect\label{section:compareBaselineBasicCount1}}
Figure~\ref{reduced:BaselineBasicCountMag1} and \ref{reduced:BaselineBasicCountDist1} show the comparison of the baseline swarm against the goal-based swarm with reduced GPS usage. Both graphs show that the settling period of the swarm from the chaotic state is very much in line with the settling period of the baseline swarm. This is due to the algorithms effect when the swarm is initially compressed. The neighbour count prevents any agents from being coordinators ($|nbr(b)| > 5$). This is supported by the propagation of the coordinator role in Figure~\ref{fig:SwarmNeighbourCount}. The swarm therefore acts in a similar manner to the baseline. As the simulation progresses the swarm settles into a condition that is similar to baseline with a higher deviation and lower mean resultant \textit{inter-agent vector}. This is due to there being a \textit{direction vector} which impacts on the coordinator agents and propagates to the non-coordinator agents through proximity. This effect is shown in Figure~\ref{metric:StabilityNullVector}.

@@

Figure~\ref{reduced:SimpleMulti12} shows a screen shot of the simulator using the basic-count algorithm. The reduction in GPS usage is limited to edge based agents of the swarm. Some perimeter agents are not identified as a coordinator due to the number of neighbours they have being $\geq5$. These agents tend to be indented on a perimeter edge. This feature will be discussed in more detail in chapter~\ref{chapter:ConcaveReduction}.

@@

The full-perimeter algorithm allows a predictable subset of agents to be identified. This subset can be used to influence the overall direction of the swarm. The full-perimeter detection algorithm detects similar agents to the basic-count algorithm but also identifies `indented' agents that the basic-count algorithm ignores. The basic-count algorithm identifies these agents as having a neighbour count beyond the threshold. These additionally detected agents are used in void reduction which is discussed in chapter~\ref{chapter:ConcaveReduction}.

@@

\section{Baseline and effect comparison\label{section:compareBaselineAll1}}
To compare the three algorithms and their effects the metric defined in chapter~\ref{chapter:metric} is combined with the swarm's emergent properties. The internal movement (jitter), the terminal speed of the centroid of the swarm and the effect on the path of the swarm must are considered.

@@

\begin{table}
\begin{center}
\begin{tabular}{| p{2.5cm} | c | c | p{5cm} |}
\hline
GPS Model & Bots GPS & \% usage & Description \\ \hline
All-agents & 200 & 100\% & All agents are coordinators \\ \hline
Basic-count & 40 & 20\% & Minimal counting algorithm\\ \hline
Full perimeter & 72 & 36\% & Void detection and perimeter (capable of supporting void reduction)\\  \hline
\end{tabular}\caption{Swarm GPS enabled coordinators} \label{tab:GPSUsage}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{| l | l | l |}
\hline
GPS Model & End & Distance \\ \hline
All-agents & (1294.5940641812, 233.788749492987) & 588.5813143506\\ \hline
Basic-count & (1201.4534856023, 289.7373900729) & 696.0606134399\\ \hline
Full Perimeter & (1218.1290381962, 260.3440891182) & 669.4489951522\\ \hline
\end{tabular}\caption[Distance to destination after run]{Distance to destination after run \textit{(end = [1841,15])}}\label{tab:SwarmDistanceEnd}
\end{center}
\end{table}

@@

\begin{table}
%\fontsize{8}{8}\selectfont
\begin{center}
\begin{tabular}{| c | r | r | r |}
\hline
{\parbox{2cm}{\centering Motor}} & {\parbox{2cm}{\centering All\\100/0}} & {\parbox{2cm}{\centering Full\\36/64}} & {\parbox{2cm}{\centering Basic \\20/80}} \\ \hline
100mA &  227.04W &  208.4544W &  203.808W\\ \hline
1A    & 2009.04W & 1990.4544W & 1985.808W\\ \hline
3A    & 5969.04W & 5950.4544W & 5945.808W\\ \hline
\end{tabular}\caption{Energy consumption of swarm} \label{tab:Energy2}
\end{center}
\end{table}

\begin{table}
%\fontsize{8}{8}\selectfont
\begin{center}
\begin{tabular}{| c | r | r | r |}
\hline
{\parbox{2cm}{\centering Motor}} & {\parbox{2cm}{\centering All\\100/0}} & {\parbox{2cm}{\centering Full\\36/64}} & {\parbox{2cm}{\centering Basic \\20/80}} \\ \hline
100mA & 0\% & 8.186\% & 10.232\% \\ \hline
1A    & 0\% & 0.925\% &  1.156\% \\ \hline
3A    & 0\% & 0.311\% &  0.389\% \\ \hline
\end{tabular}\caption{Energy consumption of swarm} \label{tab:Energy3}
\end{center}
\end{table}

@@

The SenseSwarm mechanism does not detect internal voids and is therefore not suitable for concave reduction as discussed in chapter~\ref{chapter:ConcaveReduction}.

>>>>

\begin{table}
\begin{center}
\begin{tabular}{| p{1.5cm} | c | p{3.5cm} |}
\hline
\bf Weight \bf Component & \bf Hexagonal & \bf Description \\ \hline
Sample Rate & 100 & ms - Unit sampling interval\\  \hline
$k_c$ & 5 & weight adjuster for cohesion vector\\  \hline
$k_r$ & 15 & weight adjuster for repulsion  vector\\  \hline
$k_d$ & 0/100 & weight adjuster for destination vector 0 for static baseline 100 for destination based\\  \hline
Neighbour Distance & 70 & units\\  \hline
Agent Boundary & 80 & units\\  \hline
Speed & 20 & units/s\\  \hline
\end{tabular}\caption{Swarm Weighted Model} \label{tab:DynamicPhysics1}
\end{center}
\end{table}

@@

\textit{Assumption 1:} The direction of the swarm is based upon having a fixed goal that is known to all the agents. \\
\textit{Assumption 2:} The goal is migrated towards based on a simulated GPS signal and the bias $k_d$ is applied to the \textit{destination vector} to influence the movement of the agents.

@@

\section{Identifying the coordinator role}
For an agent to change its role from a standard agent to a coordinator there needs to an algorithm that identifies a set of conditions to trigger the change. This same algorithm also triggers an agent to revert to being a standard agent. The coordinator selection algorithm is separate from the agent coordination algorithms that determine cohesion and repulsion.

@@

\section{Monolithic swarm - (all-agent)} 
In a monolithic swarm \cite{BS:13} the propagation towards the goal is achieved by all the agents using a GPS signal to give each agents' movement a \textit{direction vector} towards a required goal. There is no selection criteria and no computational overhead for the agents.

@@

\section{Simple multifaceted swarm (basic-count)} 
In a simple multifaceted swarm the propagation towards the goal is achieved by generating a \textit{destination vector} for an agent when the agent conforms to a simple counting rule. This \textit{destination vector} is obtained by using the agent's GPS to identify its current coordinate and generating a vector from that coordinate to a specific destination point. This rule creates a subset of agents that apply a local \textit{direction vector} to their movement. This has the effect of creating a directional effect on the whole swarm. The aggregate \textit{direction vector} magnitude of the swarm is less than that produced when all agents apply a \textit{direction vector}. This reduction is further diluted by agents without the \textit{direction vector} calculating a \textit{movement vector} to an equilibrium position. This change in the aggregate \textit{direction vector} will impact on a swarm's `jitter'. This is discussed in more detail in~Section~\ref{sec:AlternateBias1}. With there only being a limited number of peripheral agents invoking their GPS signal there will also be a net reduction in the energy usage of the swarm~(Figure~\ref{reduced:EnergyReductionGPS}).

A swarm with the appropriate field effects will propagate towards hexagons made from a central agent and six neighbours. A swarm can be influenced by its surroundings, which involves the interaction of other agents (as neighbours) and obstacles which can cause a compression of the agents. Compression can also occur during the stabilisation phase of a swarm. These situations can cause an agent to have less than 6 neighbours but still be surrounded. A general rule therefore to detect a rudimentary boundary of a hexagonal swarm is to calculate the number of neighbours an agent has. If the total is less than five ($nbr(b) < 5$) then there is a high probability that the agent is either in a void within the swarm or on an external perimeter. If an agent is on a perimeter or a void the it should enable its GPS and become a coordinator and provide a directional bias to the swarm via its \textit{direction vector}. This coordinator role affects the immediate neighbours \textit{movement vector} through cohesion/repulsion~(Figure~\ref{reduced:SimpleMulti12}). This simple detection algorithm allows a directional bias to be induced into a swarm with minimal computational impact on the coordinator agents.

@@

\subsection{Basic-count effect\label{section:compareBaselineBasicCount1}}
Figure~\ref{reduced:BaselineBasicCountMag1} and \ref{reduced:BaselineBasicCountDist1} show the comparison of the baseline swarm against the goal-based swarm with reduced GPS usage. Both graphs show that the settling period of the swarm from the chaotic state is very much in line with the settling period of the baseline swarm. This is due to the algorithms effect when the swarm is initially compressed. The neighbour count prevents any agents from being coordinators ($|nbr(b)| > 5$). This is supported by the propagation of the coordinator role in Figure~\ref{fig:SwarmNeighbourCount}. The swarm therefore acts in a similar manner to the baseline. As the simulation progresses the swarm settles into a condition that is similar to baseline with a higher deviation and lower mean resultant \textit{inter-agent vector}. This is due to there being a \textit{direction vector} which impacts on the coordinator agents and propagates to the non-coordinator agents through proximity.

@@

Figure~\ref{reduced:SimpleMulti12} shows a screen shot of the simulator using the basic-count algorithm. The reduction in GPS usage is limited to edge based agents of the swarm. Some perimeter agents are not identified as a coordinator due to the number of neighbours they have being $\geq5$. These agents tend to be indented on a perimeter edge.

@@

The full-perimeter algorithm allows a predictable subset of agents to be identified. This subset can be used to influence the overall direction of the swarm. The full-perimeter detection algorithm detects similar agents to the basic-count algorithm but also identifies `indented' agents that the basic-count algorithm ignores. The basic-count algorithm identifies these agents as having a neighbour count beyond the threshold.

@@

\section{Baseline and effect comparison\label{section:compareBaselineAll1}}
To compare the three algorithms and their effects the metrics applied. The internal movement (jitter), the terminal speed of the centroid of the swarm and the effect on the path of the swarm are considered.

@@

\begin{table}
\begin{center}
\begin{tabular}{| p{1.5cm} | c | c | p{3cm} |}
\hline
GPS Model & Bots GPS & \% usage & Description \\ \hline
All-agents & 200 & 100\% & All agents are coordinators \\ \hline
Basic-count & 40 & 20\% & Minimal counting algorithm\\ \hline
Full perimeter & 72 & 36\% & Void detection and perimeter (capable of supporting void reduction)\\  \hline
\end{tabular}\caption{Swarm GPS enabled coordinators} \label{tab:GPSUsage}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{| p{1.2cm} | l | l |}
\hline
GPS Model & End & Distance \\ \hline
All-agents & (1294.5940641812, 233.788749492987) & 588.5813143506\\ \hline
Basic-count & (1201.4534856023, 289.7373900729) & 696.0606134399\\ \hline
Full Perimeter & (1218.1290381962, 260.3440891182) & 669.4489951522\\ \hline
\end{tabular}\caption[Distance to destination after run]{Distance to destination after run \textit{(end = [1841,15])}}\label{tab:SwarmDistanceEnd}
\end{center}
\end{table}

@@

\begin{table}
%\fontsize{8}{8}\selectfont
\begin{center}
\begin{tabular}{| c | r | r | r |}
\hline
{\parbox{1.5cm}{\centering Motor}} & {\parbox{1cm}{\centering All\\100/0}} & {\parbox{1.5cm}{\centering Full\\36/64}} & {\parbox{1.5cm}{\centering Basic \\20/80}} \\ \hline
100mA &  227.04W &  208.4544W &  203.808W\\ \hline
1A    & 2009.04W & 1990.4544W & 1985.808W\\ \hline
3A    & 5969.04W & 5950.4544W & 5945.808W\\ \hline
\end{tabular}\caption{Energy consumption of swarm} \label{tab:Energy2}
\end{center}
\end{table}

\begin{table}
%\fontsize{8}{8}\selectfont
\begin{center}
\begin{tabular}{| c | r | r | r |}
\hline
{\parbox{1.5cm}{\centering Motor}} & {\parbox{1cm}{\centering All\\100/0}} & {\parbox{1,5cm}{\centering Full\\36/64}} & {\parbox{1.5cm}{\centering Basic \\20/80}} \\ \hline
100mA & 0\% & 8.186\% & 10.232\% \\ \hline
1A    & 0\% & 0.925\% &  1.156\% \\ \hline
3A    & 0\% & 0.311\% &  0.389\% \\ \hline
\end{tabular}\caption{Energy consumption of swarm} \label{tab:Energy3}
\end{center}
\end{table}

@@
